---
bibliography: [../../common/covidcast.bib, pnas-materials/pnas-sample.bib]
output:
  pdf_document:
    keep_tex: true
    template: pnas-suppl-template.tex
params:
  fd_casefloor: 30
  flag_jumps: 15
---


<!-- PNAS materials -->

<!-- Comment out or remove this line before generating final copy for submission; this will also remove the warning re: "Consecutive odd pages found". -->
\instructionspage



<!-- we need this -->
\maketitle

<!-- Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text. -->

\SItext

All our text goes here. Reference figures with the R chuck label as Figure~\ref{fig:fcast-finalized}.

All figures go on their own page after all of the text...

Edit `pnas-suppl-template.tex` for the correct Author list and Title.




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 6.5, fig.height = 4.5,
                      fig.align = "center",
                      fig.path = "fig/",
                      cache = TRUE,
                      out.width = "\\textwidth",
                      autodep = TRUE,
                      message = FALSE,
                      warning = FALSE)

library(covidcast)
library(gridExtra)
library(ggplot2)
library(ggforce) # for splitting across pages
library(scales)
library(dplyr)
```

\begin{figure}
\includegraphics[width=\textwidth]{fig/severity-pyramid.pdf}
\caption{The epidemiological "severity pyramid" represents the progression of
cases from the public, through infection, through increasingly severe stages of
disease. The annotations here represent the data sources collected by Delphi's
COVIDcast Epidata API. TODO Edit to use consistent language to paper}
\end{figure}

\clearpage

```{r state-trends}
SOURCES <- c("jhu-csse", "chng", "chng", "fb-survey")
SIGNALS <- c("confirmed_7dav_incidence_prop",
             "smoothed_adj_outpatient_cli",
             "smoothed_adj_outpatient_covid",
             "smoothed_whh_cmnty_cli")
colors <- c("black", scales::hue_pal()(5))
names <- c("Cases", "CHNG-CLI", "CHNG-COVID", "CTIS-CLI-in-community")

# geo_labeler is a function from geo_value to human-readable name
get_trend_grid <- function(trends, geo_labeler) {
  case_max <- max(trends[[1]]$value)

  trends <- aggregate_signals(trends, format = "long") %>%
    group_by(signal) %>%
    mutate(value = value / max(value, na.rm = TRUE) * case_max) %>%
    ungroup() %>%
    mutate(signal = factor(signal, levels = SIGNALS, labels = names),
           geo_value = geo_labeler(geo_value))

  g <- ggplot(trends, aes(x = time_value, y = value, color = signal)) +
    geom_line() +
    scale_color_manual(breaks = names, values = colors) +
    labs(x = "Date", y = "Signal value (scaled)") +
    scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
                 date_minor_breaks = "1 month", guide = guide_axis(angle = 45)) +
    theme_bw() +
    theme(legend.position = "bottom", legend.title = element_blank())

  return(g)
}

state_trends <- covidcast_signals(
  SOURCES,
  SIGNALS,
  "2020-04-15", "2021-04-15",
  geo_type = "state"
)

g <- get_trend_grid(state_trends, function(val) { abbr_to_name(toupper(val)) })
```

```{r state-trend-grids-1, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. Cases are reported cases per 100,000 population. Other signals are scaled to have the same global maximum across all counties and times, so they can be presented in the same range. (part 1 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 1)
```

\clearpage

```{r state-trend-grids-2, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 2 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 2)
```

\clearpage

```{r state-trend-grids-3, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 3 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 3)
```

\clearpage

```{r state-trend-grids-4, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 4 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 4)
```

\clearpage

```{r county-trends}
# Same thing again, but the 50 largest counties
top_counties <- county_census %>%
  filter(COUNTY != 0) %>% # these are states
  arrange(desc(POPESTIMATE2019)) %>%
  head(n = 50) %>%
  pull(FIPS)

# Work around an API problem: specifying geo_values in the query causes it to
# fail on the CHNG data because of bad database indexing. Filter after the
# fact instead. DANGER: Very slow.
county_trends_nochng <- covidcast_signals(
  SOURCES[c(1,4)], SIGNALS[c(1,4)],
  "2020-04-15", "2021-04-15",
  geo_type = "county",
  geo_values = top_counties
)

county_trends_chng <- covidcast_signals(
  SOURCES[2:3], SIGNALS[2:3],
  "2020-04-15", "2021-04-15",
  geo_type = "county"
)

county_trends_chng <- lapply(
  county_trends_chng,
  function(df) {
    filter(df, geo_value %in% top_counties)
  }
)

county_trends <- append(county_trends_nochng, county_trends_chng[1])
county_trends <- append(county_trends, county_trends_chng[2])
```

```{r county-trend-grids-1, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. Cases are reported cases per 100,000 population. Other signals are scaled to have the same global maximum across all counties and times, so they can be presented in the same range. (part 1 of 4)", fig.height=8}
g <- get_trend_grid(
  county_trends,
  function(fipsen) {
    paste0(county_fips_to_name(fipsen), ", ", name_to_abbr(state_fips_to_name(fipsen)))
  }
)

g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 1)
```

\clearpage

```{r county-trend-grids-2, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 2 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 2)
```

\clearpage

```{r county-trend-grids-3, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 3 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 3)
```

\clearpage

```{r county-trend-grids-4, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 4 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 4)
```


<!-- PNAS says: Add this line AFTER all your figures and tables -->
\FloatBarrier
